---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lnmixsurv

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/vitorcapdeville/persistencia/branch/master/graph/badge.svg)](https://app.codecov.io/gh/vitorcapdeville/persistencia?branch=master)
[![R-CMD-check](https://github.com/vivianalobo/lnmixsurv/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/vivianalobo/lnmixsurv/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `lnmixsurv` is to provide an easy interface to the bayesian lognormal mixture model described at (incluir artigo da Viviana).

An usual formula-type model is implemented in `survival_ln_mixture`, with the usual `suvival::Surv()` interface.
The model tries to follow the [conventions for R modeling packages](https://tidymodels.github.io/model-implementation-principles/),
and uses the [hardhat](https://hardhat.tidymodels.org/) structure.

The underlying algorithm implementation is a Gibbs sampler which takes initial values from a small run of the EM-Algorithm. The samplers are implemented in `C++` using `RcppArmadillo` for the linear algebra operations, `RcppGSL` for the random number generation and `OpenMP` for parallellization.

## Dependencies
Before proceeding, make sure to have [GSL](https://www.gnu.org/software/gsl/) installed on your system. Specifically, on MacOS, `brew install gsl` should be enough. On Windows/Linux, these probably came by default or together with the R install. If any errors occur, please, contact us.

## Installation
You can install the development version of `lnmixsurv` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vivianalobo/lnmixsurv")
```

## parsnip and censored extension

An extension to the models defined by [parsnip](https://parsnip.tidymodels.org/index.html) and [censored](https://censored.tidymodels.org/articles/examples.html)
is also provided, adding the `survival_ln_mixture` engine to the `parsnip::survival_reg()` model.

The following models, engines, and prediction type are available/extended through `persistencia`:

```{r, echo=FALSE, message=FALSE}
library(censored)
library(dplyr)
library(purrr)
library(tidyr)
library(lnmixsurv)

yep <- cli::symbol$tick
nope <- cli::symbol$cross
mod_names <- get_from_env("models")
model_info <-
  map_dfr(mod_names, ~ get_from_env(paste0(.x, "_predict")) %>% mutate(model = .x)) %>%
  select(model, engine, mode, type)
pkg_info <-
  map_dfr(mod_names, ~ get_from_env(paste0(.x, "_pkgs")) %>% mutate(model = .x)) %>%
  select(model, engine, pkg) %>%
  unnest(cols = pkg) %>%
  filter(pkg == "lnmixsurv") %>%
  select(!pkg)

model_info %>%
  filter(mode == "censored regression") %>%
  select(model, engine, mode, type) %>%
  pivot_wider(
    names_from = type,
    values_from = mode,
    values_fill = nope,
    values_fn = function(x) yep
  ) %>%
  inner_join(pkg_info, by = c("model", "engine")) %>%
  knitr::kable()
```
