---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lnmixsurv

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![Codecov test coverage](https://codecov.io/gh/vitorcapdeville/persistencia/branch/master/graph/badge.svg)](https://app.codecov.io/gh/vitorcapdeville/persistencia?branch=master)
[![R-CMD-check](https://github.com/vivianalobo/lnmixsurv/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/vivianalobo/lnmixsurv/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of `lnmixsurv` is to provide an easy interface to the bayesian lognormal mixture model described at (incluir artigo da Viviana).

An usual formula-type model is implemented in `survival_ln_mixture`, with the usual `suvival::Surv()` interface.
The model tries to follow the [conventions for R modeling packages](https://tidymodels.github.io/model-implementation-principles/),
and uses the [hardhat](https://hardhat.tidymodels.org/) structure.

The underlying algorithm implementation is a Gibbs sampler which takes initial values from a small run of the EM-Algorithm. The samplers are implemented in `C++` using `RcppArmadillo` for the linear algebra operations, `RcppGSL` for the random number generation and `OpenMP` for parallellization.

## Dependencies
Before proceeding, make sure to have [GSL](https://www.gnu.org/software/gsl/) and OpenMP, if using Mac OS, installed on your system. Specifically, on Mac OS, `brew install gsl` should be enough for GSL. On Windows/Linux, these probably came by default or together with the R install. If any errors occur, please, contact us.

### OpenMP -- Mac OS
The default compiler of Mac OS (clang) is shipped without support to OpenMP. Thus, installing the package directly might result in error. One option is to compile the package using **GCC**. To do so, make sure GCC is installed with `brew install gcc`. Now, create or edit the file **~/.R/Makevars** to include the lines

> CC = gcc
> 
> CPP = g++

Then, restart R session (or open one) and try to install the package again.

If the method before didn't work, it's worth trying out what is described [here](https://mac.r-project.org/openmp/). Before proceeding, make sure to have **curl** installed, `brew install curl`. Then, execute

    curl -O https://mac.r-project.org/openmp/openmp-12.0.1-darwin20-Release.tar.gz
    sudo tar fvxz openmp-12.0.1-darwin20-Release.tar.gz -C /

Now, create or edit **~/.R/Makevars** to include the lines
> CPPFLAGS += -Xclang -fopenmp
> 
> LDFLAGS += -lomp

The following piece of code can be used to do everything, all at once:

    curl -O https://mac.r-project.org/openmp/openmp-12.0.1-darwin20-Release.tar.gz && sudo tar fvxz openmp-12.0.1-darwin20-Release.tar.gz -C / && mkdir -p ~/.R && echo -e "CPPFLAGS += -Xclang -fopenmp\nLDFLAGS += -lomp" > ~/.R/Makevars

After that, the package should install with no problems at all. 

## Installation
You can install the development version of `lnmixsurv` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vivianalobo/lnmixsurv")
```

## parsnip and censored extension

An extension to the models defined by [parsnip](https://parsnip.tidymodels.org/index.html) and [censored](https://censored.tidymodels.org/articles/examples.html)
is also provided, adding the `survival_ln_mixture` engine to the `parsnip::survival_reg()` model.

The following models, engines, and prediction type are available/extended through `persistencia`:

```{r, echo=FALSE, message=FALSE}
library(censored)
library(dplyr)
library(purrr)
library(tidyr)
library(lnmixsurv)

yep <- cli::symbol$tick
nope <- cli::symbol$cross
mod_names <- get_from_env("models")
model_info <-
  map_dfr(mod_names, ~ get_from_env(paste0(.x, "_predict")) %>% mutate(model = .x)) %>%
  select(model, engine, mode, type)
pkg_info <-
  map_dfr(mod_names, ~ get_from_env(paste0(.x, "_pkgs")) %>% mutate(model = .x)) %>%
  select(model, engine, pkg) %>%
  unnest(cols = pkg) %>%
  filter(pkg == "lnmixsurv") %>%
  select(!pkg)

model_info %>%
  filter(mode == "censored regression") %>%
  select(model, engine, mode, type) %>%
  pivot_wider(
    names_from = type,
    values_from = mode,
    values_fill = nope,
    values_fn = function(x) yep
  ) %>%
  inner_join(pkg_info, by = c("model", "engine")) %>%
  knitr::kable()
```
