---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# persistencia

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

The goal of persistencia is to provide an easy interface to the bayesian lognormal mixture model described in (incluir artigo da Viviana).
An usual formula-type model is implemented in `survival_ln_mixture`, with the usual `suvival::Surv(time, event) ~ cov` interface.
The model tries to follow the [conventions for R modeling packages](https://tidymodels.github.io/model-implementation-principles/),
and uses the [hardhat](https://hardhat.tidymodels.org/) structure.

An extension to the models defined by [parsnip](https://parsnip.tidymodels.org/index.html) and [censored](https://censored.tidymodels.org/articles/examples.html)
is also provided, adding the `survival_ln_mixture` engine to the `survival_reg` model.

The underlying algorithm implementation is a gibbs sampler (incluir detalhes sobre prioris e etc) and is implemmented in `C++`, using `RcppArmadillo` for the linear algebra operations
and `OpenMP` to provide a parallelized way of generating multiple chains.

## Installation

You can install the development version of persistencia from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vitorcapdeville/persistencia")
```

## Example

The model can be used with the usual formula interface or using the `tidymodels` and `censored` structure.

```{r example}
library(persistencia)
library(dplyr)
lung$sex <- factor(lung$sex)
set.seed(1)
mod <- survival_ln_mixture(Surv(time, status == 2) ~ sex, lung, intercept = TRUE)
mod
```

For the tidymodels approach, when using `recipes`, some caution must be taken due to the fact that recipes do not support inline functions,
and the original variables can only have their roles updated with `update_role`.

```{r example_tidymodels}
library(tidymodels)
tidymodels_prefer()

dados <- lung |> mutate(outcome = Surv(time, status == 2))
rec <- recipe(outcome ~ sex, dados) |>
  step_mutate(sex = factor(sex)) |>
  step_dummy(sex)

mod_spec <- survival_reg() |>
  set_engine("survival_ln_mixture") |>
  set_mode("censored regression")

wflw <- workflow() |>
  add_model(mod_spec) |>
  add_recipe(rec)

set.seed(1)
mod_tidy <- wflw |> fit(dados)
mod_tidy
```

The predicitons are easy to obtain from a fit.

```{r prediction}
library(ggplot2)
library(purrr)
library(dplyr)
library(tidyr)

models <- list(formula = mod, tidymodels = mod_tidy)
new_data <- lung |> distinct(sex)
pred_sob <- map(models, ~ predict(.x, new_data, type = "survival", time = seq(120)))

bind_rows(pred_sob, .id = "modelo") %>%
  group_by(modelo) %>%
  dplyr::mutate(id = new_data$sex) %>%
  ungroup() %>%
  tidyr::unnest(cols = .pred) %>%
  ggplot(aes(x = .time, y = .pred_survival, col = id)) +
  geom_line() +
  theme_bw() +
  facet_wrap(~modelo)
```
