---
title: "Comparison with other models"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparison with other models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  eval = FALSE,
  include = TRUE
)
```

This vignette shows a brief comparison of the survival_ln_mixture with other survival models, 
available through `censored`.

We begin by loading the packages and preparing the data.

```{r}
library(lnmixsurv)
require(censored)
require(purrr)
require(dplyr)
require(ggplot2)

set.seed(25)

data <- simulate_data(
  n = 6000,
  mixture_components = 3,
  k = 2,
  percentage_censored = 0.3
)$data |>
  rename(x = cat, y = t)

new_data <- data |>
  distinct(x)
```

```{r eval = TRUE, include = FALSE}
library(lnmixsurv)
library(readr)
require(censored)
require(purrr)
require(dplyr)
require(ggplot2)

set.seed(25)

dir.create(file.path("data", "compare"), showWarnings = FALSE)

if (!file.exists("data/compare/data.rds")) {
  data <- simulate_data(
    n = 6000,
    mixture_components = 3,
    k = 2,
    percentage_censored = 0.3
  )$data |>
    rename(x = cat, y = t)

  write_rds(data, "data/compare/data.rds")
}


data <- read_rds("data/compare/data.rds")

new_data <- data |>
  distinct(x)

formula <- Surv(y, delta) ~ x
```

For comparison, lets also estimate the Kaplan-Meier survival function.

```{r}
require(ggsurvfit)
km <- survfit2(formula, dados)

# Manipulating the KM to format that matches parsnip predictions.
surv_km <- tidy_survfit(km, type = "surv") |>
  select(.eval_time = time, .pred_survival = estimate, id = strata) |>
  tidyr::nest(.pred = c(.eval_time, .pred_survival))
```

```{r eval = TRUE, include = FALSE}
library(ggsurvfit)

if (!file.exists("data/compare/km.rds")) {
  km <- survfit2(formula, data)

  write_rds(km, "data/compare/km.rds")
}

km <- read_rds("data/compare/km.rds")

surv_km <- tidy_survfit(km, type = "surv") |>
  select(.eval_time = time, .pred_survival = estimate, id = strata) |>
  tidyr::nest(.pred = c(.eval_time, .pred_survival))
```

The we build our `parsnip` specifications and store them in a list.

```{r eval = TRUE}
ln_survival <- survival_reg(dist = "lognormal") |>
  set_engine("survival")

ph_survival <- proportional_hazards() |>
  set_engine("survival")

decision_tree <- decision_tree(cost_complexity = 0) |>
  set_engine("rpart") |>
  set_mode("censored regression")

ln_mixture <- survival_reg() |>
  set_engine("survival_ln_mixture",
    iter = 800, warmup = 0,
    starting_seed = 5
  )

ln_mixture_em <- survival_reg() |>
  set_engine("survival_ln_mixture_em",
    iter = 150, starting_seed = 5
  )

specs <- list(
  ln_survival = ln_survival, ph_survival = ph_survival, ln_mixture = ln_mixture, decision_tree = decision_tree, ln_mixture_em = ln_mixture_em
)
```

Finally, thanks to the great parsnip API, we can fit and predict all models at once.

```{r}
set.seed(1)

models <- map(specs, ~ fit(.x, formula, dados))

pred_sob <- map(models, ~ predict(.x, new_data,
  type = "survival",
  eval_time = seq(1000)
))
```

```{r eval = TRUE, include = FALSE}
set.seed(1)

if (!file.exists("data/compare/models.rds")) {
  models <- map(specs, ~ fit(.x, formula, data))

  write_rds(models, "data/compare/models.rds")
}

models <- read_rds("data/compare/models.rds")

if (!file.exists("data/compare/pred_sob.rds")) {
  pred_sob <- map(models, ~ predict(.x, new_data,
    type = "survival",
    eval_time = seq(1000)
  ))

  write_rds(pred_sob, "data/compare/pred_sob.rds")
}

pred_sob <- read_rds("data/compare/pred_sob.rds")
```

The following plot compares each model with the Kaplan-Meier estimates of the survival function.

```{r eval = TRUE, fig.width = 7}
all_preds <- bind_rows(pred_sob, .id = "modelo") |>
  group_by(modelo) |>
  dplyr::mutate(id = new_data$x) |>
  ungroup() |>
  tidyr::unnest(cols = .pred)

km_fit <- surv_km |>
  tidyr::unnest(cols = .pred) |>
  filter(.eval_time < 1000)

ggplot(aes(x = .eval_time, y = .pred_survival, col = id), data = all_preds) +
  theme_bw() +
  geom_line() +
  facet_wrap(~modelo) +
  geom_line(data = km_fit, linetype = "dashed")
```
